Intentional release of native species undermines ecological stability

### 2022-09-28 rmd/supporting_info.Rmd
Update supporting_info.Rmd

# setup ####
source(here::here("code/library.R"))
source(here::here("code/set_functions.R"))

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

options(knitr.kable.NA = '')

# theory ####
cols <- c(
  "n_species",
  "k",
  "r_type",
  "r1",
  "r_min",
  "r_max",
  "sd_env",
  "phi",
  "int_type",
  "alpha",
  "model",
  "seed")

list_sim <- list(readRDS(here::here("output/result_ricker_2sp.rds")),
                   readRDS(here::here("output/result_ricker.rds")))

list_param <- list_sim %>% 
  lapply(FUN = function(x) {
      n_sim <- x %>% 
        mutate(n_sim = n_warmup + n_burnin + n_timestep) %>% 
        pull(n_sim) %>% 
        unique()

      param_set <- x %>% 
        dplyr::select(which(colnames(x) %in% cols)) %>% 
        distinct()

      attr(param_set, "n_sim") <- n_sim
      return(param_set)
  })

# empirical data ####

## raw data
source(here::here("code/data_fmt_fishdata.R"))
source(here::here("code/data_fmt_reg.R"))

df_area <- d0 %>% 
  group_by(site_id, year) %>% 
  summarize(area = unique(area))

## ssm estimate
df_bp <- list.files(path = here::here("output"),
                    full.names = T,
                    pattern = "summary_ssm_ar") %>% 
  readRDS() %>% 
  filter(param == "bp_value") %>% 
  dplyr::select(mean)

## mcmc sample size

df_mcmc <- lapply(c("summary_ssm_ar", "summary_reg", "summary_multi_ricker"),
       function(x) {
         list.files(path = here::here("output"),
                    full.names = T,
                    pattern = x) %>% 
           readRDS() %>% 
           ungroup() %>% 
           dplyr::select(n_total_mcmc,
                         n_sample,
                         n_thin,
                         n_burn) %>% 
           distinct() %>% 
           mutate(model = str_extract(x, "ssm|reg|ricker"))
       }) %>% 
  bind_rows()

## release data
source(here::here("code/data_fmt_stock.R"))

df_stock_stage <- df_stock %>% 
  group_by(release_stage) %>% 
  summarize(stock = sum(abundance)) %>% 
  mutate(ratio = stock / max(stock),
         prop = stock / sum(stock))

@@

**Species interaction.** We analyzed a subset of our data to estimate the strength of interspecific interactions among stream fishes. In this analysis, we focused on `r n_distinct(df_complete$site_id)` sites with (1) \> 16 years of observations and (2) no intentional release of masu salmon. These additional constraints were necessary to estimate parameters in a complex multi-species model within a state-space modeling framework.

@@

```{=tex}
\begin{align}
\varepsilon_{i,t}^{\text{state}}=\sum_{d}^{D}\zeta_{t,d}\delta_{d,i} + I_{ij}\epsilon_{i,t}
\end{align}
```
where $D$ is the number of latent factors, $\epsilon_{i,t} \sim \text{Normal}(0, \sigma_{\epsilon,i}^2)$, and $I_{ij}$ is the identity matrix ($I_{ii}=1$ and $I_{ij, i \ne j} = 0$). With this formulation, the matrix $\pmb{\Omega_{\varepsilon}}$ can be described as $\pmb{\Omega_{\varepsilon}}=\pmb{\Delta}^T\pmb{\Delta} + \text{diag}(\sigma_{\epsilon,i}^2)$, where $\pmb{\Delta}$ is the matrix of factor loadings $\delta_{d,i}$. This model structure has a merit of reduced number of parameters while capturing the correlated nature of environmental noises [@ovaskainenHowAreSpecies2017]. We set $D=2$ when the number of species in a community is greater than four, otherwise $D=1$. We reported the estimates of $\alpha_{ij}$ so the values are comparable to those used in the theoretical analysis.

@@

# Table S1
source(here::here("code/table_param_set.R"))
knitr::kable(df_param, format = "markdown")

@@

# Table S2
source(here::here("code/table_species_list.R"))

@@

# Table S6
source(here::here("code/table_prior.R"))

@@

Median estimates of the Bayesian state-space model. Site-specific parameters were excluded due to a large number of parameters.

@@

# Table S7
source(here::here("code/table_ssm.R"))

@@

# Table S8-10
## load regression tables
source(here::here("code/table_reg.R"))
j <- 7

@@

# Figure S1-8
## load figures
source(here::here("code/si_figure_theory.R"))

@@

# Figure S9-11
## load figures
source(here::here("code/si_figure_obs_dyns.R"))
j <- 8

@@

source(here::here("code/si_figure_env.R"))

@@

source(here::here("code/si_figure_cor.R"))
```

>>>>

suppressMessages(source(here::here("code/set_rmd.R")))
```

@@

**Species interaction.** We analyzed a subset of our data to estimate the strength of interspecific interactions among stream fishes. In this analysis, we focused on `r n_distinct(df_complete$site_id)` sites with $>16$ years of observations and no intentional release of masu salmon. These additional constraints were necessary to estimate parameters in a complex multi-species model within a state-space modeling framework.

@@

```{=tex}
\begin{align}
\varepsilon_{i,t}^{\text{state}}=\sum_{d}^{D}\zeta_{t,d}\delta_{d,i} + I_{ij}\epsilon_{i,t}
\end{align}
```
where $D$ is the number of latent factors, $\epsilon_{i,t} \sim \text{Normal}(0, \sigma_{\epsilon,i}^2)$, and $I_{ij}$ is the identity matrix ($I_{ii}=1$ and $I_{ij, i \ne j} = 0$). With this formulation, the matrix $\pmb{\Omega_{\varepsilon}}$ can be described as $\pmb{\Omega_{\varepsilon}}=\pmb{\Delta}^T\pmb{\Delta} + \text{diag}(\sigma_{\epsilon,i}^2)$, where $\pmb{\Delta}$ is the matrix of factor loadings $\delta_{d,i}$. This model structure has a merit of reduced number of parameters while capturing the correlated nature of environmental noises [@ovaskainenHowAreSpecies2017]. We set $D=2$ in our model. We reported the estimates of $\alpha_{ij}$ so the values are comparable to those used in the theoretical analysis.

@@

# Table S1
suppressMessages(source(here::here("code/table_param_set.R")))
knitr::kable(df_param, format = "markdown")

@@

# Table S2
suppressMessages(source(here::here("code/table_species_list.R")))

@@

# Table S6
suppressMessages(source(here::here("code/table_prior.R")))

@@

Median estimates of the Bayesian state-space model. See Durante [@duranteNoteMultiplicativeGamma2017] for multiplicative gamma priors. The parameter $\pmb{\phi}$ is a vector of prior scales ($\pmb{\phi}=2.5,...,2.5$). Site-specific parameters were excluded due to a large number of parameters.

@@

# Table S7
suppressMessages(source(here::here("code/table_ssm.R")))

@@

# Table S8-10
## load regression tables
suppressMessages(source(here::here("code/table_reg.R")))
j <- 7

@@

# Figure S1-8
## load figures
suppressMessages(source(here::here("code/si_figure_theory.R")))

@@

# Figure S9-11
## load figures
suppressMessages(source(here::here("code/si_figure_obs_dyns.R")))
j <- 8

@@

suppressMessages(source(here::here("code/si_figure_env.R")))

@@

suppressMessages(source(here::here("code/si_figure_cor.R")))
```
